---
title: "DATA 605 Week 7"
author: "Vinicio Haro"
date: "10/12/2018"
output: html_document
---

1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y .

By the definition of the CFD, we can derive the following
$$
F(y)=P(Y\le y)=1-P(Y< y)\\
=1-P(min(x_1,x_2,...x_n)>y)
$$


We know that the minimum of the xi's are greater than y when xi is greater than y for all values of i. These are i.i.d variables, we can write out the following:

$$
P(y)=1-P(x_1>y)P(x_2>y)...P(x_n>y)
$$

We consider that xi are uniformally distributed on the interval (1,k)

$$
p(x_i>y)=1-\frac{y-1}{k-1}
$$

We can now develop the distribution of y 

$$
F(y)=1-(1-\frac{y-1}{k-1})^{n}
$$


Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).

a. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years..)

$$
P(X=k)=(1-p)^{k-1}p\\
E[X]=\frac{1}{p}\\
Var[X]=\frac{1-p}{p^{2}}\\
$$
Probability of machine failure each year 
```{r}
p_fail=1/10
p_fail
```

Probability of machine not failing every year
```{r}
p_doesnot_fail<-1-p_fail
p_doesnot_fail
```

Expected value 
```{r}
e<-1/p_fail
e
```

Standard Deviation
```{r}
std <- sqrt(p_doesnot_fail/(p_fail^2))
std
```

We need to consider using the geometric to find the probability the machine will fail after 8 years. We can use some standard r functions 
```{r}
p=((1-p_fail)^(8-1))*p_fail
p
```


b. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.


$$
X\le k: P(X\le k)= e^{-\lambda x}\\
E[X]=\frac{1}{\lambda}\\
Var[x]=\frac{1}{\lambda^{2}}
$$
Probability of failing
```{r}
p_exp<-exp(-1*(8/10))
p_exp
```

probability of not failing 
```{r}
p_exp_not<-1-p_exp
p_exp_not
```

Expected Value (we can use to compute lambda)
```{r}
u=10
u
```

$$
10=\frac{1}{\lambda}\\
\lambda=.10
$$

standard deviation
```{r}
std_exp<-sqrt(1/(.10^2))
std_exp
```

c. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)

$$
P(success)=(nCk)P^{n}(1-p)^{n-k}\\
E[X]=np\\
Var[X]=np(1-p)
$$

probability of machine failure
```{r}
p_binomial<-choose(8, 0)*((.1)^(0))*(1-.1)^8
p_binomial
```

probability of non failure
```{r}
p_binomial_not<-1-p_binomial
p_binomial_not
```

Expected Value 
```{r}
e_bin<-8*.1
e_bin
```

standard deviation
```{r}
std_bin <- sqrt(8*.1*(1 - .1))
std_bin

```


d. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson.

$$
P(X=x)=\frac{\lambda^{x}e^{-\lambda}}{x!}\\
E[X]=\lambda\\
Var[X]=\lambda
$$

probability of machine failure 
```{r}
p_poi<-(exp(-1 * 10) * 10^8) / factorial(8)
p_poi
```

The expected value is 10

Standard deviation 
```{r}
e_poi<-sqrt(10)
e_poi
```












