---
title: "DATA 605 HW 10"
author: "Vinicio Haro"
date: "November 2, 2018"
output: html_document
---


Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6.
Find the probability that he wins 8 dollars before losing all of his money if:

(a) he bets 1 dollar each time (timid strategy)

What are we given? Well we know that the probability of losing is .6, therfore the probability of winning would be .4. Let p=.4 and q=.6. Let x=1 and y=8

$$
\frac{(\frac{q}{p})^{x}-1}{(\frac{q}{p})^{y}-1}

$$

$$
\frac{1-(\frac{.6}{.4})^{1}}{1-(\frac{.6}{.4})^{8}}\\
=0.02
$$
The probability comes out to roughly 0.02 

verify
```{r}
(1.5^1-1)/(1.5^8-1)
```

But what is the underlying mathematics at play? Well he only bets 1 dollar at a time so the states are 0,1,2,3,4,5,6,7 and 8. We want to show

$$
P_i=(S_8<S_0)=P(S_8<S_0|X_0=i)
$$


(b) he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy)

This method is a little more involved.We need to develop a transition matrix as follows.The states on our chain are 0, 1, 2, 4, and 8 dollars. 

$$
\begin{bmatrix} 1 & 0 & 0 & 0 & 0 \\ .6 & 0 & .4 & 0 & 0 \\ .6 & 0 & 0 & .4 & 0 \\ .6 & 0 & 0 & 0 & .4 \\ 0 & 0 & 0 & 0 & 1 \end{bmatrix}
$$


Solving this will be much more involved so we will use r. 

Initialize the system
```{r}
state_matrix <- matrix(c(1,0,0,0,0,0.6,0,0.4,0,0,0.6,0,0,0.4,0,0.6,0,0,0,0.4,0,0,0,0,1), ncol=5,nrow=5, byrow = TRUE)
state_matrix
```

create the initial state vector 
```{r}
initial <- matrix(c(0,1,0,0,0), ncol=5,nrow = 1,byrow = TRUE)
initial
```


Solve the resulting system of equations 

```{r}
p1 <- initial%*%state_matrix
p2<-p1%*%state_matrix
p3<-p2%*%state_matrix
p4<-p3%*%state_matrix
p4
```

Our probability is 0.064

(c) Which strategy gives Smith the better chance of getting out of jail?

By comparing the results of our solution vectors,the bold strategy gives Smith the best chance to get out of jail. 



